---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Linear Regression

We will predict quantitative response.

```{r}
#| echo: false
#| warning: false
#| error: false
library(ISLR)
library(tidyverse)
library(ggthemes)
library(sjPlot)
library(corrplot)
```

```{r}
#| warning: false
#| error: false
advertising = read_csv("./data/Advertising.csv") %>% as_tibble %>% select(-1)
advertising
```

We are asked to suggest a marketing plan for next year which will yeild high product sales. We may want to inquire the following questions:

1.  *Is there a relationship between advertising budget and sales?*

    First we should determine if there is an association between adveritisng expenditure and sales. If not, no money shpuld be spent on advertising.

2.  *How strong is the relationship between advertising budget and sales?*

    If there is a relationship between advertising and sales, what is the strength of this relationship? Given a certain advertising budget, can we predict sales with a high level of accuracy? =\> strong relationsihp.

3.  *Which media contribute to sales?*

    Do all variables--tv,radio,newspaper-- contribute to sales, or just one or the two?

4.  *How accurately can we estimate the effect of each medium on sales?*

    For every ollar spent on advertising in a particular medium, by what amount will sales increase? How accuretly can we predict this amount of increase?

5.  *How accurately can we predict future sales?*

    For any given level of media advertisig, what is our prediction for sales, and what is the accuracy of this prediciton?

6.  *Is the relationship linear?*

    If so linear regresion is appropriate tool, if not we may need to transform the predictor or the repsonse so that liner regression can be used.

7.  *Is there synergy among the advertising media?*

    Does the effect of a medium on sales depend on other medium levels? Does dividing advertisement budget to two or three medium yeild a higher sales?

We can answer each of these questions using Linear regression.

## Simple Linear Regression

Predicting a quantitative response $Y$ on the basis of a single predictor variable $X$.

Our assumption is that there is approximately a linear relationship between $X$ and $Y$; we can write this linear relationship as

$$
Y = \beta_0 + \beta_1 X_1 + \epsilon
$$

$$
Y \approx \beta_0 + \beta_1X
$$ (3.1)

For example lets say $X$ is `TV`, and $Y$ is `sales`

$$
sales = \beta_0 + \beta_1 \times TV + \epsilon
$$ or

$$
sales = \beta_0 + \beta_1 \times TV
$$

On (3.1) $\beta_0$ and $\beta_1$ are unknown constants that represent the *intercept* and *slope* in the linear model. Together they are known as *coefficients* or *parameters*.

We are going to use or training data to produce estimates for $\beta_0$ =\> $\hat{\beta_0}$ and $\beta_1$ =\> $\hat{\beta_1}$. Using these predicted coefficients we can predict sales;

$$
\hat{sales} = \hat{\beta_0} + \hat{\beta_1} \times TV
$$

or as in general form

$$
\hat{y} = \hat{\beta_0} + \hat{\beta_1}x
$$ (3.2)

### Estimating the coefficients

Since, $\beta_0$ and $\beta_1$ are unknown, before we can use (3.1) to make predictions we must use data to estimate the coefficients. We have $n$ observations :

$$
(x_1,y_1), (x_2,y_2), \dots, (x_n,y_n)
$$

We want our estimated coefficients to give such predictions that will fit the avaible data as well =\> $y_i \approx \hat{\beta_0} + \hat{\beta_1}x_i$ for $i = 1,\dots, n$. This coefficients will allow us to draw a regression line and we want this regression line to be close as possible to the $n$ data points we have.

There are different ways to measure *closeness*. The most common approach is minimizing the *least squares* criterion. Alternative approaches will be considered in Chapter 6.

Our predictions come from $\hat{y_i} = \hat{\beta_0} + \hat{\beta_1}x_i$.

Then for each data we have a *residual*: difference between $y$ and $\hat{y}$:

$$
e_i = y_i - \hat{y_i}
$$

We need to take the squares to get the distances--because of the negative residuals, and sum them to get the *residual sum of squares*(RSS)

$$
RSS = e_1^2 + e_2^2 + \dots + e_n^2
$$

this is equal to

$$
RSS = (y_i - \hat{beta_0} - \hat{\beta_1}x_1)^2 + (y_2 - \hat{\beta_0} - \hat{\beta_1}x_2) + \dots + (y_n - \hat{\beta_0} - \hat{\beta_1}x_n)
$$ (3.3)

The least squares approach chooses $\hat{\beta_0}$ and $\hat{\beta_1}$ to minimize RSS. These minimizers are

$$
\begin{align}
\hat{\beta_1} &= \frac{\sum_{i=1}^n(x_i - \bar{x})(y_i - \bar{y})}{\sum_{i=1}^n(x_i - \bar{x})^2} \\
\hat{\beta_0} &= \bar{y_i} - \hat{\beta_1}\bar{x}
\end{align}
$$ (3.4)

$\bar{y} = \frac{1}{n}\sum_{i=1}^ny_i$ and $\bar{x} = \frac{1}{n}\sum_{i=1}^nx_i$ are the sample means.

So (3.4) defines the *least squares coefficient estimates* for simple linear regression.


Lets calculate them with R
```{r}
beta_1_hat_adv = sum((advertising$TV - mean(advertising$TV)) * ((advertising$sales - mean(advertising$sales)))) / sum((advertising$TV - mean(advertising$TV))^2)
beta_1_hat_adv
```

So; our $\hat{\beta_1} = 0.0475 $ 
```{r}
beta_0_hat_adv = mean(advertising$sales) - beta_1_hat_adv * mean(advertising$TV)
beta_0_hat_adv
```
our $\hat{\beta_0} = 7.032$

Lets compare them with r function
```{r}
summary(lm(sales ~ TV, data = advertising))
```


Calculating the predicted values 
```{r}
y_hat_adv = beta_0_hat_adv + beta_1_hat_adv * advertising$TV
y_hat_adv[1:10]
```
Our $y_i$ values are as above.

```{r}
RSS = sum((advertising$sales - y_hat_adv)^2)
RSS
```

Our $\text{RSS} = 2102.531$

So we can draw our regression line

```{r fig.width=10}
#| warning: false
#| error: false
#| fig-cap: For the advertising data, the least squares fit for the regression of sales onto TV. The fit is found by minimizing the sum of squared errors. Each grey line segment represents an error, adn the fit make a comprimise by averaging their squares. In this case a linear fit captures the essence of the relationship, although it is somewhat deficient in the left of the plot
advertising %>%
  ggplot() + aes(x=TV, y = sales) +  geom_abline(intercept = beta_0_hat_adv, slope = beta_1_hat_adv, color = "#262B70", size =1.2) +
  geom_segment(aes(xend=TV, yend=y_hat_adv), color = "#939393") +
  geom_point(color = "#AA1D2E", size =2) + theme_par()
```


So we have

$$
\hat{y_i} = 7.032 + 0.0475x_i
$$

According to this approximation an additional \$1,000 spent on TV increases sales by 47.5 units.

### Assessing the Accuracy of the Coefficient Estimates

We assumed that *true* relationship is linear: $Y = f(X) + \epsilon$. We don't know $f$, and $\epsilon$ is a mean-zero random error term.

We said $f$ is approximatly linear, so that $f(X) = \beta_0 + \beta_1 X$; which means

$$
Y = \beta_0 + \beta_1 X + \epsilon
$$
(3.5)

error term captures: 
* the true relationship may not be linear
* other variables that affect Y
* measurement error

and is independent of $X$.


(3.5) is the *population regression line*: the best linar approximation to the true relationship between $X$ and $Y$.

$$
\hat{y} = \hat{\beta_0} + \hat{\beta_1}X
$$
is the *least squares line*. They are different of course! But we don't know the population regression line. If we did:

For example, lets create a data;

* First we create random x values from 100 random numbers

```{r}
seq(-2,2,length.out=100)[1:10]
```

```{r}
set.seed(11)
x = sample(seq(-2,2,length.out = 100),size = 100, replace = T)
x[1:10]
```

lets define our $f$--population parameters $\beta_0$ and $\beta_1$

$$
f(X) = 2 + 3\times X
$$
Now lets create our $Y$ values from this function but we also want to add random error values as well
```{r}
set.seed(11)
y = 2 + 3*x + rnorm(100)
y[1:10]
```

So we have a data set

```{r}
data = tibble(
  y = y, x = x
)
```

Now lets plot this data points

```{r fig.width=8}
data %>% 
  ggplot() + aes(x=x, y=y) + geom_point() + theme_par()
```

Now, even though we already know $f$ and population parameters $\beta_0 = 2$ and $\beta_1 = 3$, lets estimate them:

```{r}
summary(lm(y ~ x, data = data))
```

So our *least squares estimation is*

$$
\hat{y_i} = 1.88 + 3.06 x_i
$$
Lets draw this *least squares regression line to our plot*

```{r}
data %>% 
  ggplot() + aes(x=x, y=y) + geom_point() + geom_abline(intercept=1.88, slope = 3.06, size = 1.2) + theme_par()
```

What about the population regression line

```{r}
data %>% 
  ggplot() + aes(x=x, y=y) + geom_point() + geom_abline(intercept = 1.9, slope = 3.06, size = 1.2) + geom_abline(intercept = 2, slope = 3, color ="red") + theme_par()
```

They are not the same! If we were to have another data from the same data generation process other estimates of parameters would result with different *least squares regression lines*:

```{r}
set.seed(111)
x_r1 = sample(seq(-2,2,length.out = 100),size = 100, replace = T)
y_r1 = 2 + 3*x_r1 + rnorm(100)
set.seed(1111)
x_r2 = sample(seq(-2,2,length.out = 100),size = 100, replace = T)
y_r2 = 2 + 3*x_r2 + rnorm(100)
set.seed(11111)
x_r3 = sample(seq(-2,2,length.out = 100),size = 100, replace = T)
y_r3 = 2 + 3*x_r3 + rnorm(100)
set.seed(111111)
x_r4 = sample(seq(-2,2,length.out = 100),size = 100, replace = T)
y_r4 = 2 + 3*x_r4 + rnorm(100)
set.seed(1111111)
x_r5 = sample(seq(-2,2,length.out = 100),size = 100, replace = T)
y_r5 = 2 + 3*x_r5 + rnorm(100)
set.seed(11111111)
x_r6 = sample(seq(-2,2,length.out = 100),size = 100, replace = T)
y_r6 = 2 + 3*x_r6 + rnorm(100)
```

Lets now estimate population parameters for each of these data and plot them

```{r}
data_r = tibble(
  y_r1,x_r1,y_r2,x_r2,y_r3,x_r3,y_r4,x_r4, y_r5,x_r5,y_r6,x_r6
)
data_r
```

```{r}
data %>% 
  ggplot() + aes(x,y) + geom_point(size = 0, color = "white")  + 
  geom_abline(intercept = lm(y_r1 ~ x_r1)$coefficients[1], slope = lm(y_r1 ~ x_r1)$coefficients[2], color ="#29019F") +
  geom_abline(intercept = lm(y_r2 ~ x_r2)$coefficients[1], slope = lm(y_r2 ~ x_r2)$coefficients[2], color ="#0A04BF") +
  geom_abline(intercept = lm(y_r3 ~ x_r3)$coefficients[1], slope = lm(y_r3 ~ x_r3)$coefficients[2], color ="#0930DF") +
  geom_abline(intercept = lm(y_r4 ~ x_r4)$coefficients[1], slope = lm(y_r4 ~ x_r4)$coefficients[2], color ="#0E6DFF") +
  geom_abline(intercept = lm(y_r5 ~ x_r5)$coefficients[1], slope = lm(y_r5 ~ x_r5)$coefficients[2], color ="#2BA8FF") +
  geom_abline(intercept = lm(y_r6 ~ x_r6)$coefficients[1], slope = lm(y_r6 ~ x_r6)$coefficients[2], color ="#48D9FF") +
  geom_abline(intercept =2, slope =3, color = "red") +
  
  theme_par()
```


So, different data sets generated from the same true model result in slightly different least squares lines, but the unobserved population regression line does not change.

This is because we are using a sample, and estimating characteristics of the population. Usually these characteristics are different, but generally sample characteristics will provide a good estimate to the population characteristics.

Computing $\hat{\beta_0}$ and $\hat{\beta_1}$ from different sets of sample data provide different but similar results. And we are trying to estimate population parameters $\beta_0$ and $\beta_1$ with these. Some of these $\hat{\beta_0}$ and $\hat{\beta_1}$ will overestimate, some will underestimate $\beta_0$, and $\beta_1$. But if we could average all these estimated parameters and take the average, than this average should be equal to population parameters; if this is the case this estimator is called *unbiased estimator*. So an unbiased estimator does not *systematically* over- or under-estimate the true parameter.

Okay but how close $\hat{\beta_0}$ and $\hat{\beta_1}$ are to the true values $\beta_0$ and $\beta_1$. We want to compute the standard errors associated with $\hat{\beta_0}$ and $\hat{\beta_1}$. Standard error telss us the average amount of estimate differes from the actual value.

$$
\begin{align}
\text{SE}(\hat{\beta_0})^2 &= \sigma^2 \left[\frac{1}{n} + \frac{\bar{x}^2}{\sum_{i=1}^n(x_i - \bar{x}^2)}\right] \\
\text{SE}(\hat{\beta_1})^2 &= \frac{\sigma^2}{\sum_{i=1}^n(x_i-\bar{x})^2}
\end{align}
$$
(3.8)

Where $\sigma^2=\text{Var}(\epsilon)$.

Notice that formula of $\text{SE}(\hat{\beta_1})$ is smaller when $x_i$ are more spread out; intutively we have more *leverage* to estimate a slope when this is the case.

In general $\sigma^2$ is not known, but can be estimated from the data. The estimate of $\sigma$ is known as the *residual standard error*, and given by the formula

$$
\text{RSE} = \sqrt{\text{RSS}/(n-2)}
$$
So, when $\sigma^2$ is estimated fro mthe data we should write $\hat{\text{SE}}(\hat{\beta_1})$ to indicate that an estimate has been made, but usually we drop this extra hat.

Standard errors can be used to compute *confidence intervals*. A 95% confidence interval is defines as a range of values such that with 95% probability, the rage will contain the true unknown value of the parameter. The range is defined in terms of lower and upper limits computed from the sample of data. For linear regression, the 95% confidence interval for $\beta_1$ approximately takes the form

$$
\hat{\beta_1} \pm 1.96 \cdot \text{SE}(\hat{\beta_1}) 
$$
(3.9)

So there is approximately a 95% chance that the interval 
$$[\hat{\beta_1} - 1.96 \cdot \text{SE}(\hat{\beta_1}), \hat{\beta_1} + 1.96 \cdot \text{SE}(\hat{\beta-1})]$$ 
(3.10) will contain the true value of $\beta_1$. Same is true for $\beta_0$

$$
\hat{\beta_0} \pm 1.96 \cdot \text{SE}(\hat{\beta_0})
$$
(3.11)


Lets calculate the confidence intervals for $\beta_0$ and $\beta_1$ from our original data and model

$$
\hat{sales_i} = \hat{\beta_0} + \hat{\beta_1}\cdot TV
$$
We first need to calculate RSS and RSE:

$$
\text{RSS} = \sum(y_i - \hat{y_i})^2
$$

```{r}
RSS = sum((advertising$sales - y_hat_adv)^2)
RSS
```

$$
\text{RSE} = \sigma = \sqrt{RSS/(n-2)}
$$
```{r}
RSE = sqrt((RSS / (length(advertising$sales) -2)))
RSE
```

For $\beta_0$

$$
\text{SE}(\hat{\beta_0}) = \sigma^2 \left[\frac{1}{n} + \frac{\bar{x}^2}{\sum_{i=1}^n(x_i - \bar{x})^2} \right]
$$

```{r}
se_beta_0_adv =  sqrt(RSE^2 * (1/length(advertising$sales) + (mean(advertising$TV)^2 / sum((advertising$TV - mean(advertising$TV))^2))))
se_beta_0_adv
```

So we can calculate the confidence interval for $\beta_0$

```{r}
cat("In the absence of any advertising, sales will on average, fall somewhere between",beta_0_hat_adv - 1.96 * se_beta_0_adv, "and", beta_0_hat_adv + 1.96 * se_beta_0_adv)
```

For $\hat{\beta_1}$:

$$
\text{SE}(\hat{\beta_1})^2 =\frac{\sigma^2}{\sum(x_i - \bar{x})^2}
$$

```{r}
se_beta_1_adv = sqrt(RSE^2 / (sum((advertising$TV - mean(advertising$TV))^2)))
se_beta_1_adv
```

```{r}
cat("For each $1,000 increase in TV advertising, average increase in sales will be between",(beta_1_hat_adv - 1.96 * se_beta_1_adv) * 1000, "and", (beta_1_hat_adv + 1.96 * se_beta_1_adv)*1000, "by 95% confidence")
```



Lets confirm our results

```{r}
summary(lm(sales ~ TV, advertising))
```

```{r}
confint(lm(sales ~ TV, advertising))
```

So standard errors of our estimated parameters tells us the average amount of difference from the true population parameters. And using the confidence intervals we can tell a range of the true population parameters' interval with a percentage (usually 95%).

Lets do this for our `data` as well, which we know has the form

$$
y_i = 2 + 3 x_i + \epsilon_i
$$
Lets calculate $\hat{\beta_0}$ and $\hat{\beta_1}$ first

```{r}
beta_1_hat_data = sum((data$x - mean(data$x)) * (data$y - mean(data$y))) / sum((data$x - mean(data$x))^2)
beta_1_hat_data
```

```{r}
beta_0_hat_data = mean(data$y) - beta_1_hat_data * mean(data$x)
beta_0_hat_data
```

$$
\hat{y_i} = 1.88 + 3.05 x_i
$$

lets confirm this 

```{r}
lm(y~x, data)
```

Lets calculate the residual sum of squares, residual sum of errors, standard errors, and confidence intervals

```{r}
data %>% 
  mutate(
    y_hat = beta_0_hat_data + beta_1_hat_data * x
  ) -> data

```

```{r}
RSS_data = sum((data$y - data$y_hat)^2)
RSS_data
```



```{r}
RSE_data = sqrt((RSS_data/(length(data$y) -2)))
RSE_data
```

so $\sigma_{data} = 0.9163$

We can now compute the standard errors of estiamed coefficients

```{r}
se_beta_0_data = sqrt(((1/length(data$y)) + (mean(data$x)^2 / sum((data$x - mean(data$x))^2))) * RSE_data^2)
se_beta_0_data
```

```{r}
se_beta_1_data = sqrt(RSE_data^2 / (sum((data$x - mean(data$x))^2)))
se_beta_1_data
```

Lets confirm
```{r}
summary(lm(y~x,data))
```

So we can say that on average our $\hat{\beta_0}$s are 0.091 differ from $\beta_0$, and our $\hat{\beta_1}$s differ 0.076 from $\beta_1$. To make more sense of it we can calculate the confidence intervals

```{r}
cat("By 95% confidence we can say that the true beta_0 is between", beta_0_hat_data - 1.96 * se_beta_0_data, "and", beta_0_hat_data + 1.96 * se_beta_0_data )
```

```{r}
cat("By 95% confidence we can say that the true beta_0 is between", beta_1_hat_data - 1.96 * se_beta_1_data, "and", beta_1_hat_data + 1.96 * se_beta_1_data)
```

Lets confirm this

```{r}
confint(lm(y~x,data))
```

Since standard error tells us the range of the $\beta$ values via confidence interval, we can infer that if this range does not include 0, than our $\beta$ values are statistically significant; x is assocaited with y.

Or we can use standard erros to perform *hypothesis tests* on the coefficients. We usually don't care about the intercept, so lets do the hypothesis test on only $\hat{\beta_1}$.

$$
\begin{align}
H_0 &: \beta_1 = 0 \to \text{there is no relationship between X and Y} \\
H_1 &: \beta_1 \neq 0 \to \text{there is some relationship between X and Y}
\end{align}
$$
If the *null-hypothesis* is true => $ \beta_1 = 0$ => $Y = \beta_0 + \epsilon$ => $X$ is not associated with $Y$.

To test the null-hypothessi, we need to determine whether our estimate $\hat{\beta_1}$ is sufficiently far from zero that we can be confident that $\beta_1$ is non-zero. How far is enough? This depends on the accuracy of $\hat{\beta_1}$--that is it depends on $\text{SE}(\hat{\beta_1})$. If $\text{SE}(\hat{\beta_1})$ is small, then even relatively small values of $\hat{\beta_1}$ may provide strong evidence that $\beta_1 \neq 0$. If $\text{SE}(\hat{\beta_1})$ is large, then $\hat{\beta_1}$ must be large in absolute value in order for us to reject the null hypothessis. In practice we compute a *t-statistic* given by

$$
t = \frac{\hat{\beta_1} - 0}{\text{SE}(\hat{\beta_1})}
$$
(3.14)

which measures the number of standard deviations that $\hat{\beta_1}$ is away from zero. From the t-statistic we can compute the *p-value*; a small p value indicates that it is unlikely to observe such a substantial association between the predictor and the response due to chance, in absence of any real association between the predictor and the response. So if p value is small we infer that there is assocaition between the predictor and the response => we reject the null hypothesis. Typical p-value cutoffs for rejecting the null hypothesis are 5 or 1%. When $n=30$ these correspond to tstatsitcs of around 2 and 2.75.

```{r}
summary(lm(sales ~ TV, advertising))
```

Here we see that t statistics are very high, and p values are very low => reject the null hypothesis for both $\beta$ values; they are statistically significant.

### Assessing the Accuracy of the Model

Once we concluded the statistically significant variable--rejecting the null hypothesis, we want to quantify *the extend to which the model fits the data*. We can use either

* *Residual standard error*
* $R^2$

*Residual standard error*

Recall from $Y_i = \beta_0 + \beta_1 X_i + \epsilon_i$ that associated with each observation is an error term $\epsilon$. Because of these error terms even if we knew the true regression line, we would not be able to predict $Y$ from $X$. The *RSE* is an estimate of the standard deviation of $\epsilon$. It is the average amount that the response will deviate from the true regression line, computed by

$$
\begin{align}
\text{RSE} &= \sqrt{\frac{1}{n-2}\text{RSS}} \\
&= \sqrt{\frac{1}{n-2}\sum_{i=1}^n(y_i - \hat{y_i})^2}
\end{align}
$$
(3.15)

In the advertising data, RSE was 3.26; actual sales in each market deviate from the true regression line by approximately 3,269 units, on average. This also means that; if the model were correct and the true values of the unknown coefficients $\beta_0$ and $\beta_1$ were known exaclty, any predcition of sales on the basis of TV advertising would still be off by about 3,260 units on average. Is this prediction error accaptable? Depends on the data: in the `advertising` data set the mean value of `sales` is $\approx 14,000$ units, and so the percentage error is $3,260 / 14,000 = 23%$.

The RSE is considered a measure of the *lack of fit* of the model $Y=\beta_0 + \beta_1 + \epsilon$ to the data. If the predictions from the model are very close to the true outcome values--$\hat{y_i} \approx y_i$ then RSE will be small, and we can concldue that the model fits the data very well. Otherwise, if $\hat{y_i}$ is very far from $y_i$ then RSE may be quite large, indicating the model doesn't fit the data well.

$R^2$**Statistic**

The RSE provides an absolute measure of lack of fit of the model to the data. $R^2$ provides an alternative measure of fit. It takes the form of a *proportion*--the proportion of variance explained--and so its always $0\leq R^2 \leq 1$ and is independent of the scale of $Y$--as opposed to RSE.

$$
R^2 = \frac{TSS - RSS}{TSS} = 1 - \frac{RSS}{TSS}
$$
(3.17)

where $\text{TSS} = \sum(y_i - \bar{y})^2$ is the *total sun of squares* and $\text{RSS} = \sum(y_i - \hat{y_i})^2$. TSS measures the total variaance in the response $Y$; and can be thought of as the amount of varaiblity ingerent in the response before the regression is performed. RSS measures the amount of varaiblity that is left unexplained after performing the regression. So TSS - RSS measures the amount of variability in the response that is explained by performing the regression, and $R^2$ measures the *proportion of variability in* $Y$ *that can be explained using* $X$. As $R^2$ gets closer to 1, a large proportion of the variability in the response has been explained by the regression. A number near 0 indicates that the regression did not explain much of the variablity in the response; this might occur because the linear model is wrong, or the inherit error $\sigma^2 = \text{RSE}^2$ is high, or both.

Lets calculate $R^2$ of our estimation on `advertising` data with the model $\hat{sales_i} = \hat{\beta_0} + \hat{\beta_1}TV_i$

```{r}
advertising %>% 
  mutate(sales_hat = beta_0_hat_adv + beta_1_hat_adv * TV) -> advertising
advertising
```

```{r}
RSS = sum((advertising$sales - advertising$sales_hat)^2)
RSS
```

```{r}
RSE = sqrt(RSS/(length(advertising$sales) -2))
RSE
```

```{r}
TSS = sum((advertising$sales - mean(advertising$sales))^2)
TSS
```

```{r}
R2 = (TSS - RSS) / TSS
R2
```

61% of the variablity in `sales` is explained by a linear regression on `TV`.

what about our `data`

```{r}
data
```

```{r}
RSS_data = sum((data$y - data$y_hat)^2)
RSS_data
```

```{r}
RSE_data = sqrt(RSS_data/(length(data$y) -2))
RSE_data
```

```{r}
TSS_data = sum((data$y - mean(data$y))^2)
TSS_data
```

```{r}
R2_data = (TSS_data - RSS_data) / TSS_data
R2_data
```

94% of the variablity in `y` is explained by `x`; very good fit of the model.

$R^2$ is better to interpret than RSE.

## Multiple Linear Regression

In practice we have more than one predictor to explain $Y$. 

How can we extend our analysis of the advertising order to accomodate the other two (`radio` and `newspaper`) additional predictors?

=> We can run three separate simple linear regressions, each of which uses a different advertising medium as a predictor:

```{r}
summary(lm(sales ~ TV, advertising))
```

```{r}
summary(lm(sales ~ radio, advertising))
```

```{r}
summary(lm(sales ~ newspaper, advertising))
```

We find that on average, \$1,000 increase in spending on radio advertising is associated with an increase in sales by around 203 units.

We find that on average, \$1,000 increase in spending on newspaper advertising is associated with an increase in sales by around 55 units.

We find that on average, \$1,000 increase in spending on TV advertising is associated with an increase in sales by around 47 units.


**However** this approach is not good. First of all it is unclear to make a sinlge prediction of sales given levesl of the three advertising media budgets, since each has their own regression equation. Second, each of these three regression equations ignores the  other two medi in forming estimates for the regression coefficients. Especially if these media budgets are correalted, this can lead to very misleading estimates of the individaul media effects on sales.

Instead of the seperate linear regressions for each predictor, better approach is to extend the simple linear regression setting $Y = \beta_0 + \beta_1 X$ to

$$
Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \dots + \beta_p X_p + \epsilon
$$

(3.19)

Here we interpret $B_j$ as the average effect of a one unit increase in $X_j$, *holding all other predictors fixed*.

For the advertising example;

$$
sales = \beta_0 + \beta_1 TV + \beta_2 radio + \beta_3 newspaper + \epsilon
$$

### Estimating the Regression Coefficients

Again, regression coefficients in (3.19) are unknown, and must be estimated from the data. And with these estimates we can make predictions

$$
\hat{y_i} = \hat{\beta_0} + \hat{\beta_1}x_1 + \hat{\beta_2}x_2 + \dots + \hat{\beta_p}x_p
$$
(3.21)

The parameters are estimated using the same least squares approach with simple linear regression. We choose $\beta_0, \beta_1, \dots, \beta_p$ to minimize the sum of squared residuals

$$
\begin{align}

\text{RSS} &= \sum_{i = 1}^n(y_i - \hat{y_i})^2 \\
&= \sum_{i = 1}^n(y_i - \hat{\beta_0} - \hat{\beta_1}x_{i1} - \beta_2x_{i2} - \dots - \beta_px_{ip})^2

\end{align}
$$
(3.22)

$\hat{\beta_0}, \hat{\beta_1},\dots, \hat{\beta_p}$ values minimize RSS.

We are not going to calculate these estimates with our hands, R does that.

Lets see our model results with the three predictors.


```{r}
summary(lm(sales ~ TV + radio + newspaper, advertising))
```

*Interpretation:* for a given amount of Tv and newspaper advertising, spending additional \$1,000 on radio(TV)(newspaper) advertising leads to an increase in sales approximately by 189(46)(-1) units.

If we compare these effects with one predictor regressions

> We find that on average, \$1,000 increase in spending on radio advertising is associated with an increase in sales by around 203 units.

> We find that on average, \$1,000 increase in spending on newspaper advertising is associated with an increase in sales by around 55 units.

> We find that on average, \$1,000 increase in spending on TV advertising is associated with an increase in sales by around 47 units.

For tv and radio coefficients are similar, but for **newspaper**: from the simple linear regression coefficient of newspaper was significant, but in multiple linear regression it is not; p value is very high.

```{r}
confint(lm(sales ~ TV + radio + newspaper, advertising))
```
Its confidence interval contains 0.

This difference between simple linear regression and multiple linear regression coefficients stems from the fact that in the simple regression, the slope term represents the average effet of a one dollar increase in newspaper advertising, ignoring other preditors such as tv and radio. In contrsat, in the multiple regression setting, the coefficient for newspaper represents the average effect of increeasing newapper spending by one dollar, while holding tv and radio fixed.

Does it make sense for the multiple regression to suggest no relationship between sales and newspaper while the simple linear regression implies the opposite? Yes! 

Take a look at this correlation matrix:
```{r}
cor(advertising[1:4])
```

We can also make it a plot out of this:

```{r}
corrplot(cor(advertising[1:4]), method = "number")
```


Notice that correlation between radio and newspaper is 0.35. This reveals a tendency to spend more on newspaper advertising in markets where more is spent on radio advertising. Now suppose the multiple regression is correct and newspaper advertising has no direct impact on sales, but radio advertising does increase sales. Then in markets where we spend more on radio, our sales will tend to be higher, adn as our correaltion matrix shows, we also tend to spend more on newspaper advertising in those same markets. Hence, in a simple linaer regresion which only examines sales vs newspaper, we will observe that higher values of newspaper tend to be associated with higher values of sales, even though newspaper advertising does not actually affect sales. So newspaper sales are proxy for radio advertising; newspaper gets credit for the effect of radio on sales.

This is a very common issue. Consider running a regression of shark attack versus ice cream sales for data collected at a given beach community. We would see a positive relationship, similar to that seen between sales and newspaper. Of course ice creams doesnt cause shark attacks. In reality higher temperatures cause more people to visit the beach in trun results in more ice cream sales and more shark attacks. A multiple regression of attacks versus ice cream sales and temperature revals that, the former predictr is no longer significant after adjusting for temperature.

### Some important Questions

When we perform MLR, we usually are interested answering a few important questions

1. *Is at least one of the predictors* $x_1, x_2, \dots, x_p$ *useful in predicting the response?*

2. *Do all predictors help to explain* $Y$, *or is only a subset of the predictors useful?*

3. *How well does the model fit the data?*

4 *Given a set of predictor values, what response value should we predict, and how accurate is our prediction?*

Lets answer these questions:

1. *Is at least one of the predictors* $x_1, x_2, \dots, x_p$ *useful in predicting the response?*

    In SLR we simply checked whether $\beta_1 = 0$  or not. In MLR, we need to ask whether all of the regression coefficients are zero $\beta_1 = \beta_2 = \dots = \beta_p = 0$. So our null hypothesis is
    
    $$
    \begin{align}
    H_0 &: \beta_1 = \beta_2 = \dots = \beta_o = 0 \\
    H_\alpha &: \text{at least one} \space B_j \space \text{is non-zero}
    \end{align}
    $$
    This hypothesis test is performed by computing the *F-statistic*,
    
    $$
    F = \frac{(TSS - RSS)/p}{RSS/(n-p-1)}
    $$
    (3.23)
    
    So, if there is no relationship between the resposne and predictors, we expect F-statistic to take on value close to 1. if $H_\alpha$ is true then $F$ should be greater than 1.
    
```{r}
summary(lm(sales ~ TV + radio + newspaper, advertising))
```


   F statistic is 570 and is far from 1. But it is best to have a look at the p-value of the F statistic which is also very small.
    
   This means that at least one of the media is associated with increase sales.
    
  In (3.23) we are testing $H_0$ that all the coefficients are zero. Sometimes we want to test that a particular subset of $q$ of the coefficients are zero. This corresponesd to a null hypothesis
    
  $$
  H_0 : \beta_{p-q+1} = \beta_{p-q+2} = \dots = \beta_p
  $$
  
  In this case we fit a second model that uses all the variables *except* those last $q$. suppose that residual sum of squares for that model is $RSS_0$. Then the appropriate F-statistic is
  
  $$
  F = \frac{(RSS_0 - RSS)/q}{RSS/(n-p-1)}
  $$
  (3.24)
  
  
  On advertising MLR we saw that newspaper is not significant from its p value. Then why do we need to look at the overall F-statistic? after all, it seems likely that if any one of the p-values for the individual variables is very small, then *at least one of the predictors is realted to the respose*. This is not true usually, especially when $p$ is large.

  So after estimating the model first look at the F-statistic, than to the individual t statistic p values.
  
2. *Do all predictors help to explain* $Y$, *or is only a subset of the predictors useful?* => **Deciding on important variables**

    After lookig at the F statistic, we can look at the individual p values. But if *p$ is large, we are going to make false discoveries. 
    
    Usually not all predictors are associated with the response. This task of determining which predictors are associated with the response in order to fit a single model involving only those predictors is refered to as *variable selection*. Check out Chapter 6 for more detail. But here is a breif outline of some of the classical approaches.
    
    Ideally we want to perform variable selection by trying out a lot of different models, each containing different subset of the predictors. For instance if our $p=2$ then we can consider four models
    
    * (1) a model containing no variables
    * (2) a model containing $x_1$ only
    * (3) a model containnig $x_2$ only
    * (4) a model containing $x1$ and $x_2$.
    
    We can then select the *best* model out of all the models by looking at some statistics we can use to judge the quality of the model. These are
    
    * *Mallow*'s $C_p$
    * *Akaike information creterion* (AIC)
    * *Bayesian information criterion*(BIC)
    * *adjusted* $R^2$
    
    These are discussed in more detail in chapter 6.
    
    We can also determine which model is the best by plotting various model outputs, such as the residuals, in order to search for patterns.
    
    But we cannot consider all models, especially when $p$ is high. There are three classical approaches for this task:
    
    * *Forward selection* 
        * begin with *null model* a model that contains an intercept but no predictors.
        
        * Then fit *p* simple linear regressions and add to the null model the variable that results in the lowest RSS. 
        
        * Then add to that model the variable that results in the lowest RSS for the new two-variable model. This approach is continued until some stopping rule is satisfied.
      
    * *Backward selection*
      * Put all varaibles in the model.
      * remove the least statistically significant predictor.
      * estimate the new regression with $p-1$ variable, remove the largest p-value predictor. This procedure continues until a stopping rule is reached => stop after all remaining variables have p value < 0.02
      
    * *Mixed selection*
      * Combination of forward selection and backward selection
      * Start with no variables in the model
      * add the varaible that provides the best fit
      * add varaibles one-by-one
      * at one point if the p-value for one of the variables in the model rises above a certain treshold, then we remove that variabel from the model.
      * Continue untill all variables have sufficiently low p value, and all vairables in the model woudl have a large p-value if added to the model
      
    
    Backwar slecetion cannot be used if $p>n$, forward selection can always be used. 
    
3. *How well does the model fit the data?* **Model Fit**

Two of the most common numerical measures of model fit are RSE and $R^2$.

In SLR $R^2$ is equal to $cor(Y,X)$. In MLR $R^2 = cor(Y,\hat{Y})$.

$R^2$ will always increase as you add more variable, even though that varaible is not statistically significant. This is because adding another variable must allow us to fit the trainig data(not necessarly test data) more accurately. But this increase in $R^2$ after adding a statistically not-significant varible is very low => evidence that you can drop the not significant variable. 
Check out the $R^2$ variables of the following models

```{r}
summary(lm(sales ~ TV + radio + newspaper, advertising))
```

```{r}
summary(lm(sales ~ TV + radio, advertising))
```

They are almost the same.

But lets see the $R^2$ of the model containing only Tv

```{r}
summary(lm(sales ~ TV, advertising))
```

it is 0.611.

If we add radio

```{r}
summary(lm(sales ~ TV + radio, advertising))
```

It increaes dramatically. This implies that model that uses TV and radio to predict sales is better than only using Tv. also radio is statistically signifiacnt.

```{r}
summary(lm(sales ~ TV + newspaper, advertising))
```

Not with  newspaper though.

Or the opposite

```{r}
summary(lm(sales ~ radio, advertising))
```

```{r}
summary(lm(sales~radio+TV, advertising))
```

```{r}
summary(lm(sales ~ radio + newspaper, advertising))
```

What about RSE:

```{r}
summary(lm(sales ~ TV + radio, advertising))
```

```{r}
summary(lm(sales ~ TV + radio + newspaper, advertising))
```

adding newspaper increased the RSE => no need to add newspaper. Adding newspaper increases RSE because

$$
RSE = \sqrt{\frac{1}{n-p-1}RSS}
$$

Models with more variables can have higher RSE if the decrease in RSS is small relative to the increase in $p$.

So we can look at both the RSE and $R^2$.




        

